Using Sweat Na/Cl Ratio to Predict CFTR-RD and CRMS

### 2024-03-25 Edith Zemanick/Sweat Test/sweat_test_cf_conference.qmd
CF research conference background

---
title: "Using Sweat Na/Cl Ratio to Predict CFTR-RD and CRMS"
author: "Tim Vigers"
date: "today"
date-format: long
format: revealjs
editor: source
bibliography: Presentation.bib
csl: /Users/timvigers/GitHub/styles/american-medical-association.csl
---

@@

@@

---
nocite: |
  @jamesIntroductionStatisticalLearning2021, @robinPROCOpensourcePackage2011, @therneau1997introduction
---

>>>>

---
title: "Using Sweat Na/Cl Ratio to Predict CFTR-RD and CRMS"
author: "Tim Vigers"
date: "today"
date-format: long
format: revealjs
editor: source
bibliography: Presentation.bib
csl: /Users/timvigers/GitHub/styles/american-medical-association-10th-edition.csl
---

@@

## Sweat chloride testing

- Sweat chloride testing is the gold standard for diagnosing Cystic Fibrosis (CF).

| Chloride level (mmol/L) | Diagnosis                |
|-------------------------|--------------------------|
| >= 60                   | Cystic Fibrosis          |
| 30 - 59                 | Intermediate diagnosis   |
| < 30                    | Cystic Fibrosis unlikely |

## Intermediate sweat test

- Can be caused by:
  - CF
  - CF Carrier
  - CFTR-Related Metabolic Syndrome (CRMS)
    - Cystic Fibrosis Screen Positive, Inconclusive Diagnosis (CFSPID)
  - CFTR-related disorder (CFTR-RD)

## CRMS/CFSPID

- Inconclusive CF diagnosis following newborn screen
- Elevated IRT, intermediate sweat, and less than 2 disease-causing CFTR variants
- No clinical signs of Cystic Fibrosis
- Should be followed by CF provider to monitor for symptoms of CF @bombieriRecommendationsClassificationDiseases2011
  - Close follow-up, repeat sweat testing, and CFTR sequencing in the first months of life 
  - Then annual follow-up until age 6-7

## CFTR-RD

- Cystic Fibrosis Transmembrane Regulator (CFTR)-related disorders
  - "A clinical entity associated with CFTR dysfunction that does not fulfill diagnostic criteria for CF" @bombieriRecommendationsClassificationDiseases2011
  - Has symptoms, but no positive newborn screen, sweat test, and less than 2 disease-causing CFTR variants
- Should be followed closely by PCP with consultation of a CF provider to monitor for symptoms of CF

## Using sodium and chloride for diagnosis

- Small studies have shown that 
  - The ratio of sweat sodium to chloride (Na:Cl) is significantly different between pwCF and non-CF patients @hallSweatSodiumChloride1990
  - Suggested cutoffs are anywhere from 1.0-1.2 @hallSweatSodiumChloride1990 @treggiariRoleSweatIon2021
- This ratio may be more helpful in diagnosing patients with intermediate sweat chloride values @treggiariRoleSweatIon2021
- *Can we use both sodium and chloride to better predict CFTR-RD/CRMS?*

@@

---
nocite: |
  @jamesIntroductionStatisticalLearning2021, @robinPROCOpensourcePackage2011, @therneau1997introduction, @metcalf_utility_2024
---
